---
title: "Introduction: a simple demo"
author: "Paul Shannon"
package: rnaSeqNormalizer
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{"Introduction: a simple demo"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
.main-container { width: 1200px; max-width:2800px;}
</style>


```{r setup, include = FALSE}
options(width=120)
knitr::opts_chunk$set(
   collapse = TRUE,
   eval=TRUE,
   echo=TRUE,
   comment = "#>"
)
```

# Overview

RNA-seq data comes from many sources in many forms.  The goal of this package is to collect
a full bag of tricks and techniques to transform that variety into our (that is, Cory Funk's)
preferred standard form.

# Load the libraries we need

```{r loadLibraries,  echo=TRUE, results=FALSE, message=FALSE}
library(rnaSeqNormalizer)
library(RUnit)
```

Read in a small tsv file included in this package

```{r readFile, results='show'}
file <- system.file(package="rnaSeqNormalizer", "extdata", "tbl.ensg.column.16x10.tsv")
tbl.small <- read.table(file, sep="\t", as.is=TRUE)
checkEquals(dim(tbl.small), c(16, 10))
head(tbl.small)
````

Use the "asinh" algorithm and use maximum median value to select among duplicated gene symbols
sometimes produced in the mapping from ENSG ids.

```{r normalize, results='show'}
normalizer <- rnaSeqNormalizer(tbl.small, algorithm="asinh", duplicate.selection.statistic="median")
mtx.asinh.median <- getNormalizedMatrix(normalizer)
fivenum(mtx.asinh.median)
head(mtx.asinh.median)
hist(as.numeric(mtx.asinh.median))
```



# Session Info

```{r sessionInfo}
sessionInfo()
````

