---
title: "Introduction: a simple demo"
author: "Paul Shannon"
package: rnaSeqNormalizer
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{"Introduction: a simple demo"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
.main-container { width: 1200px; max-width:2800px;}
</style>


```{r setup, include = FALSE}
options(width=120)
knitr::opts_chunk$set(
   collapse = TRUE,
   eval=interactive(),
   echo=TRUE,
   comment = "#>"
)
```


# Overview

RNA-seq data comes from many sources in many forms.  The goal of this package is to collect
a full bag of tricks and techniques to transform that variety into our (that is, Cory Funk's)
preferred standard form.

# Load the libraries we need

```{r loadLibraries,  results='hide'}
library(rnaSeqNormalizer)
library(RUnit)
```

Create an instance of the normalizer, supplying it with a small example expression matrix.  We expect
that it will have gene identifiers as row names and sample identifiers as column names.

```{r create, results='hide'}
mtx <- get(load(system.file(package="rnaSeqNormalization", "extdata", "mtx.mayoTcx.100x300.RData")))
normalizer <- rnaSeqNormalization(mtx)
```

Invoke the normalization

```{r run, results='show'}
mtx.norm <- normalize(normalizer)
checkEqualsNumeric(fivenum(mtx.norm), c(-3.59, -0.62, 0.028, 0.66, 3.23), tol=1e-2)
checkEquals(dim(mtx), dim(mtx.norm))
checkEquals(rownames(mtx), rownames(mtx.norm))
checkEquals(colnames(mtx), colnames(mtx.norm))
```



# Session Info

```{r sessionInfo}
sessionInfo()
````

