---
title: "Normalize AMP-AD Expression Files From Sage"
author: "Paul Shannon"
package: rnaSeqNormalizer
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{"Sage AMP-AD Files"}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
.main-container { width: 1200px; max-width:2800px;}
</style>


```{r setup, include = FALSE}
options(width=120)
knitr::opts_chunk$set(
   collapse = TRUE,
   eval=TRUE,
   echo=TRUE,
   comment = "#>"
)
```


# Overview

Sage provides RNA-seq expression files from the AMP-AD consortium.  They appear to be
in a standard format: tab-delimited, 4 summary lines, Ensembl identifiers in the first
column.

This vignette shows how to
  - read a file into a data.frame
  - perform some minor transformations on that data.frame to make it ready for
    rnaSeqNormalizer
  - choose parameters and perform the desired normalization
  - do some visual exploration of the results

# Load the libraries we need

```{r loadLibraries,  results=FALSE, message=FALSE}
library(rnaSeqNormalizer)
```

Create an instance of the normalizer, supplying it with a small example expression matrix.  We expect
that it will have gene identifiers as row names and sample identifiers as column names.

*Note* that the file used here, _ROSMAP_all_counts_matrix.txt_ is very large, and is *not* included
in the package.   This example therefore, works verbatim only on my build system.  To reproduce
this chunk of code, you must provide your own filepath variable.

```{r create, results='hide'}
filepath <- "~/github/rnaSeqNormalizer/inst/extdata/sage/ROSMAP_all_counts_matrix.txt"
stopifnot(file.exists(filepath))
tbl.mt <- read.table(filepath, sep="\t", as.is=TRUE, header=TRUE, nrow=-1)
tbl.mt <- tbl.mt[-c(1:4),]   # remove first four lines, which have summary statistics
colnames(tbl.mt)[1] <- "ensembl_id"   # give the first column the name we expect
tbl.mt$ensembl_id <- sub("\\.[0-9]+", "", tbl.mt$ensembl_id)   # remove version suffixes from the ids
```

We demonstrate two normalizations.  The results are quite different.  First the "log+scale"
normalization proposed by Max Robinson and Michael Wainberg.

```{r log+scale, results='hide'}
x <- rnaSeqNormalizer(tbl.mt, algorithm="log+scale", duplicate.selection.statistic="mean")
mtx.1 <-  getNormalizedMatrix(x)
fivenum(mtx.1)
```

Now use a standard Bioconductor algorithm using the DESeq package, *vst* (variance stabilization transformation)
```{r vst, results='hide'}
x <- rnaSeqNormalizer(tbl.mt, algorithm="vst", duplicate.selection.statistic="sd")
mtx.2 <-  getNormalizedMatrix(x)
fivenum(mtx.2)
```

Compare these two techniques:
```{r hist1, results='show'}
hist(as.numeric(mtx.1), main="log+scale")
```
```{r hist2, results='show'}
hist(as.numeric(mtx.2), main="vst")
```


# Session Info

```{r sessionInfo}
sessionInfo()
````

